### 2014-10-09 PrivacyProtection-2014-09-09.tex
Updating Legal Aspects

\section{Histrocial Approaches to Privacy}

\subsection{Aristotle}

Public Sphere vs. Private Sphere of politics. This distinction is concerned with the politic (public) life of politicians in early democracies opposed by their domestic (private) life.

\subsection{Warren and Brandeis}

The Right to Privacy, 1890: Establish the term Informational Privacy. They argue innovations like photography and a growing press create the need for a new right to privacy, "the right to be left alone".

The term informational privacy as conceptual foundation of privacy originates from the juristic discussion The Right to Privacy (1890) by Warren and Brandeis.
The arise of new technologies like photography combined with a growing press created an increased publicity.
Warren and Brandeis were worried that existing law like castle doctrines or libel and slander could not suffice.
They felt a need for a new right to "protect the extent to which one's thoughts, sentiments, and emotions could be shared with others" \cite{1}.
This right would be "the right to be left alone".

The right to "protect the extent to which one's thoughts, sentiments, and emotions could be shared with others" by Warren and Brandeis was later extend to cover arbitrary information and is now kown as the privacy concept "... the control we have over information about ourselves ..." \cite{2}.

\subsection{Fried}

C. Freid, An Anatomy of Values, Harvard Univ. Press, 1970

Defines privacy as the ability to control information about oneself. He argues that limited knowledge does not automatically create privacy.

@@

\subsection{European Convention on Human Rights - Article 8}

``(1) Everyone has the right to respect for his private and family life, his home and his correspondence. (2) There shall be no interference by a public authority with the exercise of this right except such as is in accordance with the law and is necessary in a democratic society in the interests of national security, public safety or the economic well-being of the country, for the prevention of disorder or crime, for the protection of health or morals, or for the protection of the rights and freedoms of others.'' \cite{2}

``The European Court of Human has given this article a very broad interpretation in its jurisprudence.'' \cite{2}

@@

\emph{Directive 95/46/EC}

EU directives in general do not hold any direct legal binding for EU citizens.
Member states have to create their own legal implementation.

``The directive regulates the processing of personal data regardless of whether such processing is automated or not.'' \cite{1} where \emph{personal data} is ``any information relating to an identified or identifiable natural person [...] (art. 2a)'' and \emph{processing} is ``any operation or set of operations which is performed upon personal data [...] (art. 2b)'' \cite{1}

The ``identified or identifiable natural person'' is called \emph{data subject}.

The sole responsibility for compliance is held by so called \textbf{controllers}.
A controller is an actor ``which alone or jointly with others determines the purposes and means of the processing of personal data; (art 2d)''
\cite{1}

This rule is applicable ``whenever the controller uses equipment situated within the EU''\cite{1}.
This will hold for every e-commerce provider \textbf{inside or outside} of the EU because the customer's computer is \emph{situated within the EU} anyways.

``Personal data should not be processed at all, except when certain conditions are met'' \cite{1}.
These conditions incorporate the seven OECD principals and are further categorized in the categories:

\begin{itemize}
\item
\textbf{Transparency}: ``The data subject has the right to be informed when his personal data is being processed [...] (art. 10 and 11)'' \cite{1} and he has the right how and by whom it is processed.
Additionally other specified requirements have to be met, i.e.~giving consent.
\item
\textbf{Legitimate purpose}: ``Personal data can only be processed for specified explicit and legitimate purposes and may not be processes further in a way incompatible with those purposes (art. 6b)''
\cite{1}.
\item
\textbf{Proportionality}: ``Personal data may be processed only insofar as it is adequate [...] in relation to [its] purposes [...]; every reasonable step must be taken to ensure that data which are inaccurate or incomplete [...] are erased or rectified (art. 6)'' \cite{1}
\end{itemize}

>>>>

% \section{Histrocial Approaches to Privacy}

@@

\subsection{European Convention on Human Rights}

The \emph{Europen Convetion on Human Rights} \cite{ECHR} was created by the members of the Europen Council in 1953 as part of the aftermath of the second world war.
All member states of the Council of Europe, which includes all EU members, have incorporated this convention into their national law.

Article 8 of the ECHR contains a protection of personal data:

\begin{quote}
(1) Everyone has the right to respect for his private and family life, his home and his correspondence.

(2) There shall be no interference by a public authority with the exercise of this right except such as is in accordance with the law and is necessary in a democratic society in the interests of national security, public safety or the economic well-being of the country, for the prevention of disorder or crime, for the protection of health or morals, or for the protection of the rights and freedoms of others.''
\end{quote}

The protection of personal data under article 8 is not an absolute law but must be considered in relation to other laws such as the freedom of expression granted in Article 10 of the ECHR.
The \emph{European Court of Human Rights} (ECtHR) overseas the implementation of this directive and has adopted a rather broad interpretation of the article.\footnote{\url{http://en.wikipedia.org/wiki/Article_8_of_the_European_Convention_on_Human_Rights}}

@@

The \emph{Directive 95/46/EC} of 1995 \cite{DIR95} is the fundamental retulation of data protection in the Europen Union.
EU directives in general do not hold any direct legal binding for EU citizens, but has to be implemented by national law of the EU memberstates.

It roughly follows the OECD guidelines for privacy protection issues in 1980 \cite{OECD80}.

The directive regulates the processing of personal data regardless of whether such processing is automated or not.
Where \emph{personal data} is any information relating to an identified or identifiable natural person [...] (Art. 2a)
and \emph{processing} is any operation or set of operations which is performed upon personal data [...] (Art. 2b).

The identified or identifiable natural person is called \emph{data subject}.

The sole responsibility for compliance is held by so called \emph{controllers}.
A controller is an actor which alone or jointly with others determines the purposes and means of the processing of personal data (Art 2d).

This rule is applicable whenever the controller uses equipment situated within the EU.
This will hold for every e-commerce provider inside or outside of the EU because the customer's computer is situated within the EU anyways.

Personal data should not be processed at all, except when certain conditions are met.
These conditions incorporate the seven OECD principals and are further categorized in the categories:

\begin{itemize}
\item \textbf{Transparency}  (\cite{1} art. 10,11): ``The data subject has the right to be informed when his personal data is being processed [...]'' and he has the right how and by whom it is processed.
Additionally other specified requirements have to be met, i.e.~giving consent.
\item \textbf{Legitimate purpose} (\cite{1} art. 6b): ``Personal data can only be processed for specified explicit and legitimate purposes and may not be processes further in a way incompatible with those purposes ''.
\item \textbf{Proportionality}: ``Personal data may be processed only insofar as it is adequate [...] in relation to [its] purposes [...]; every reasonable step must be taken to ensure that data which are inaccurate or incomplete [...] are erased or rectified (art. 6)''
\end{itemize}
